Improving \textsl{Gaia} parallax precision \\ with a data-driven model of stars

### 2018-07-30 paper/paper.tex
final referee report ;

The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
stellar parallax measurements.
Only a small fraction (but large number) of these
measurements will be precise and purely astrometric
\added{The long-term complete distance information coming from \gaia\ will
involve astrometric, photometric, and spectral information, and the balance
of these contributions will be a strong function of stellar brightness (\citealt{apsis}).}
These spectrophotometric models, along with \gaia's on-board
low-resolution $BP/RP$ spectrophotometry, can be used to provide
more precise parallax estimates for stars with poor astrometric parallax data.
The full stack required for these parallax inferences is complex.
It involves modeling the stars, as well as the dust, in the Milky Way,
and the response of the telescope itself.
It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.

@@

Here we build and use just such a data-driven model to infer more precise parallaxes for each star.
In particular, we use all of the \gaia\ \tgas\ data (\citealt{tgas};
these data were generated by the methodology of \citealt{michalik15}), that match to the
\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\psone; \citealt{ps1})
to make use of the \cite{green15} dust map. We use this data to make a model of the
noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
We build the \cmd\ using an Empirical-Bayes approach known as Extreme Deconvolution (XD;
\citealt{bovy11}).
This method deconvolves heteroskedastic data to derive an
estimate of the noise-free or low-noise distribution that \emph{would
  have} been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
perform quasar target selection
in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
\added{It also handles incomplete data, in the sense of there being missing
data values (these are like data values with arbitrarily large uncertainties);
this aspect of XD is not so relevant to this project,
where we have an essentially complete catalog.
(There is another sense for the word ``incomplete'' in this area, which has to do with
the completeness or selection function of the survey as a whole. We address this
kind of completeness in our assumption called ``selection'' above and in the Discussion.)}
XD's disadvantage, for the present purposes, is that it requires
or implies a strictly Gaussian noise model.
As we show below, this requires us to transform the \cmd\ to a space in which
the noise is approximately Gaussian.

@@

Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
about the Milky Way, and similarly weak assumptions about the properties of
stars.
To date, there has been some discussion in the literature
about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
These papers use a \gaia\ measurement to construct a parallax likelihood,
and combine this with a parallax prior.
While this is sensible and correct, all present methods proposed along
these lines build in informative (and known-to-be-wrong) assumptions
about the line-of-sight parallax distribution to \gaia\ stars.
That is, they build in strong and informative assumptions about the Milky
Way.
The \added{fundamental and critical} assumption made in this work is that for every star
in \tgas, there are other, photometrically (and bolometrically)
similar stars.
That is a much weaker assumption than has been made in most other Bayesian
distance estimations.

@@

To infer the true parallax for a star using the noise-deconvolved \cmd\ as
a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
  and will also be different than local
  disk stars in age and metallicity. However, the stationarity assumption is fairly weak:
  the requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[selection] We have no model for the selection function of \tgas\ (but see \cite{bovy17}),
  nor selection volumes for any kinds of stars.
  For this reason, we are building a model of \emph{the contents of \tgas}, not the
  properties of the stars in the Milky Way (nor any volume-limited component or region
  thereof). For this reason, our \cmd\ prior might not be appropriate to use with other surveys.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical-Bayes, or maximum-marginalized-likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar---for example, a white dwarf---inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
  (the \texttt{parallax{\_}error} entry in the \tgas\ catalog), are correct.
  Here we are making use of the fact that when
  de-noising data, under estimating measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise. \added{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior. We also ignore the star to star covariances because they are not available in the \tgas\ dataset.}
\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred parallax. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \added{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
\begin{equation}
y_0 = \varpi\,10^{\frac{1}{5}\,J^{\corr}_n} = 10^{\frac{1}{5}\,M^{\corr}_{J,n} + 2}
\label{eq:transform}
\end{equation}
where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
The transformation of the (dust corrected) absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
\added{This transformation also allows our method to handle negative parallaxes from the \tgas\ dataset naturally. A negative parallax simply refers to a distant object that is affected by large errors, and therefore intrinsically bright, and our transformation places these objects in the bright portion of the \cmd.}
Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
\begin{equation}
\begin{aligned}
\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K_s)^{\true}_n] \\
\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J-K_s)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
\mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
\mathrm{and \; color} \quad &(J-K_s)^{\corr}_n = J_n - K_{s,n} - Q_{JK}\,E(B-V)_n \\
\mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
\end{aligned}
\label{eq:data}
\end{equation}
where $E(B-V)_n$ is the output of the dust model of \cite{green15}, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.

@@

We use stars crossmatched in \tgas\ and \tmass.
The match was done using a nearest-neighbor algorithm with a search radius of $4$ arcsec\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}.
\added{There is also a \tgas\-\tmass\ cross-match as part of the official \gaia\ product, but this project began before that catalogue was available.}
We also required that the stars lie within the
observing footprint of \psone\ to access the \cite{green15} 3D dust model.
We require that the photometry have real values, and nonzero, real errors, and
remove a small selection of \tmass\ stars that have zero color and zero $J$-band
magnitude.
The full data set is visualized in the \cmd\ in \figurename~\ref{fig:data}.
The left panel shows the point estimates of the color and absolute magnitude,
using the point estimate of the parallax from \tgas. The $1\sigma$ and
$2\sigma$ contours of the distribution are shown in grey.
The right panel shows a subset of the data with the associated error bar for each star.
The uncertainties in the colors are fairly well behaved, but the large
uncertainties in some absolute magnitudes are due to the large uncertainties in
the parallax measurements.

@@

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior_odd.pdf}
\caption{\added{Five example posterior PDFs over parallax using our method for more rare objects.
The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented as a 2D histogram. In blue, we show the posterior expectation value and $1 \sigma$ uncertainties. 
We chose example points that lie outside the well populated portions of the \cmd\ to illustrate how our method works for rare \tgas\ objects. 
The remaining panels show the posterior PDF in blue, the prior PDF as shaded in teal, and likelihood of the data in black. 
Examples 2 and 4 are not very affected by the prior (teal), where good data is good data.
Examples 1, 3, and 5 show a more precise posterior PDF compared with the likelihood. For these examples, the posterior PDF is very similar to the prior PDF and might be biasing the posterior PDF estimate due to a lack of prior support at those values. These might be places where our stationarity assumption breaks down.}}
\label{fig:posterior_odd}
\end{figure}

@@

We select sources from our catalog within a $0.75~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical orange dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined parallax of the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{m67.pdf}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $0.75~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in teal.
The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the orange dotted lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in teal, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more narrow and clustered than the \tgas\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}

\subsection{Distances from Asteroseismology}
\added{As a second comparison of these photometric parallax inferences, we compare to the inferred distances to the stars from asteroseismology.  Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars than the photometric parallax method. There are about $2200$ stars that overlap our data set here and the asteroseimic data from \citealt{huber17}. In \figurename~\ref{fig:seismic} we compare the estimates in the parallax space. We chose the parallax space, rather than the distance space, since the asteroseismic distance estimates are more precise, making it a less skewed inverted pdf. We compare the cumulative difference in the parallax estimates in units of mas; the same units as the precision of \tgas. The comparison shows that our photo parallax estimates are more similar to the asteroseismic estimates than the \tgas\ likelihoods. The left panel and the middle panel show the one to one comparison of asteroseismic parallax estimates to the \tgas\ likelihoods (in black) and the photo parallax posterior pdf (in blue) respectively for all $2200$ stars. The right panel shows the cumulative distribution function of the differences between the asteroseismic estimates and the \tgas\ likelihoods (in black) or the photoparallax estimates (in blue) for each star. The comparison to the photo parallax estimates, represented by the blue line, shows a distribution with smaller differences, and therefore represents estimates that are more similar to the asteroseismic estimates.}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{photoVSseismic.pdf}
\caption{\added{Proof of concept: parallax estimate comparisons. All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate (in black) or our photo parallax estimate (in blue). And the figure is meant to guide visual comparisons between \tgas\ and photo parallax, and therefore highlight differences in these estimates. The first panel shows the one to one comparison of the \tgas\ parallax estimates and the inverted distance estimates from \citealt{huber17} using asteroseismology, and the associated $1\sigma$ uncertainties. The orange line shows the one to one line. The second panel is similar, but instead of the \tgas\ parallax estimates, it visualizes our photo parallax estimates. This shows the shrinkage of the precision of the parallax estimates, and that the expectation values are more in agreement with the asteroseismic estimates. The third panel panel visualizes the difference between the asteroseismic estimate and a \gaia\ estimate, \tgas\ in grey and photo parallax in blue, as a cumulative distribution function of the difference in units of mas. The CDF of the photo parallax estimates is shifted to smaller differences, so the photoparallax estimates are therefore more similar to the asteroseismic estimates than \tgas\ likelihoods alone.}}

@@

\section{Discussion}

@@

\section{Conclusion}
We forgo an explicit conclusion; we attempt to summarize the conclusions in the Abstract above.

>>>>

The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
stellar parallax measurements.
Only a small fraction (but large number) of these
measurements will be precise and purely astrometric.
\edit1{The long-term complete distance information coming from \gaia\ will
involve astrometric, photometric, and spectral information, and the balance
of these contributions will be a strong function of stellar brightness (\citealt{apsis}).}
These spectrophotometric models, along with \gaia's on-board
low-resolution $BP/RP$ spectrophotometry, can be used to provide
more precise parallax estimates for stars with poor astrometric parallax data.
The full stack required for these parallax inferences is complex.
It involves modeling the stars, as well as the dust, in the Milky Way,
and the response of the telescope itself.
It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.

@@

Here we build and use just such a data-driven model to infer more precise parallaxes for each star.
In particular, we use all of the \gaia\ \tgas\ data (\citealt{tgas};
these data were generated by the methodology of \citealt{michalik15}), that match to the
\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\psone; \citealt{ps1})
to make use of the \cite{green15} dust map. We use this data to make a model of the
noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
We build the \cmd\ using an Empirical-Bayes approach known as Extreme Deconvolution (XD;
\citealt{bovy11}).
This method deconvolves heteroskedastic data to derive an
estimate of the noise-free or low-noise distribution that \emph{would
  have} been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
perform quasar target selection
in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
\edit1{It also handles incomplete data, in the sense of there being missing
data values (these are like data values with arbitrarily large uncertainties);
this aspect of XD is not so relevant to this project,
where we have an essentially complete catalog.
(There is another sense for the word ``incomplete'' in this area, which has to do with
the completeness or selection function of the survey as a whole. We address this
kind of completeness in our assumption called ``selection'' above and in the Discussion.)}
XD's disadvantage, for the present purposes, is that it requires
or implies a strictly Gaussian noise model.
As we show below, this requires us to transform the \cmd\ to a space in which
the noise is approximately Gaussian.

@@

Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
about the Milky Way, and similarly weak assumptions about the properties of
stars.
To date, there has been some discussion in the literature
about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
These papers use a \gaia\ measurement to construct a parallax likelihood,
and combine this with a parallax prior.
While this is sensible and correct, all present methods proposed along
these lines build in informative (and known-to-be-wrong) assumptions
about the line-of-sight parallax distribution to \gaia\ stars.
That is, they build in strong and informative assumptions about the Milky
Way.
The \edit1{fundamental and critical} assumption made in this work is that for every star
in \tgas, there are other, photometrically (and bolometrically)
similar stars.
That is a much weaker assumption than has been made in most other Bayesian
distance estimations.

@@

To infer the true parallax for a star using the noise-deconvolved \cmd\ as
a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
  and will also be different than local
  disk stars in age and metallicity. However, the stationarity assumption is fairly weak:
  the requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[selection] We have no model for the selection function of \tgas\ (but see \cite{bovy17}),
  nor selection volumes for any kinds of stars.
  For this reason, we are building a model of \emph{the contents of \tgas}, not the
  properties of the stars in the Milky Way (nor any volume-limited component or region
  thereof). For this reason, our \cmd\ prior might not be appropriate to use with other surveys.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical-Bayes, or maximum-marginalized-likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar---for example, a white dwarf---inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
  (the \texttt{parallax{\_}error} entry in the \tgas\ catalog), are correct.
  Here we are making use of the fact that when
  de-noising data, under estimating measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise. \edit1{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior. We also ignore the star to star covariances. \deleted{because they are not available in the \tgas\ dataset}.}
\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred parallax. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \edit1{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
\begin{equation}
y_0 = \varpi\,10^{\frac{1}{5}\,J^{\corr}_n} = 10^{\frac{1}{5}\,M^{\corr}_{J,n} + 2}
\label{eq:transform}
\end{equation}
where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
The transformation of the (dust corrected) absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
\edit1{This transformation also allows our method to handle negative parallaxes from the \tgas\ dataset naturally. A negative parallax simply refers to a distant object that is affected by large errors, and therefore intrinsically bright, and our transformation places these objects in the bright portion of the \cmd.}
Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
\begin{equation}
\begin{aligned}
\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K_s)^{\true}_n] \\
\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J-K_s)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
\mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
\mathrm{and \; color} \quad &(J-K_s)^{\corr}_n = J_n - K_{s,n} - Q_{JK}\,E(B-V)_n \\
\mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
\end{aligned}
\label{eq:data}
\end{equation}
where $E(B-V)_n$ is the output of the dust model of \cite{green15}, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.

@@

We use stars crossmatched in \tgas\ and \tmass.
The match was done using a nearest-neighbor algorithm with a search radius of $4$ arcsec\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}.
\edit1{There is also a \tgas\-\tmass\ cross-match as part of the official \gaia\ product \cite{Marrese17}, but this project began before that catalogue was available.}
We also required that the stars lie within the
observing footprint of \psone\ to access the \cite{green15} 3D dust model.
We require that the photometry have real values, and nonzero, real errors, and
remove a small selection of \tmass\ stars that have zero color and zero $J$-band
magnitude.
The full data set is visualized in the \cmd\ in \figurename~\ref{fig:data}.
The left panel shows the point estimates of the color and absolute magnitude,
using the point estimate of the parallax from \tgas. The $1\sigma$ and
$2\sigma$ contours of the distribution are shown in grey.
The right panel shows a subset of the data with the associated error bar for each star.
The uncertainties in the colors are fairly well behaved, but the large
uncertainties in some absolute magnitudes are due to the large uncertainties in
the parallax measurements.

@@

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior_odd.pdf}
\caption{\edit1{Five example posterior PDFs over parallax using our method for more rare objects.
The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented as a 2D histogram. In blue, we show the posterior expectation value and $1 \sigma$ uncertainties. 
We chose example points that lie outside the well populated portions of the \cmd\ to illustrate how our method works for rare \tgas\ objects. 
The remaining panels show the posterior PDF in blue, the prior PDF as shaded in teal, and likelihood of the data in black. 
Examples 2 and 4 are not very affected by the prior (teal), where good data is good data.
Examples 1, 3, and 5 show a more precise posterior PDF compared with the likelihood. For these examples, the posterior PDF is very similar to the prior PDF and might be biasing the posterior PDF estimate due to a lack of prior support at those values. These might be places where our stationarity assumption breaks down.}}
\label{fig:posterior_odd}
\end{figure}

@@

We select sources from our catalog within a $0.75~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
\edit2{Due to the small number of M67 members in the \tgas\ dataset, the reliability of this test is slightly suspect, but still interesting.}
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical orange dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined parallax of the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{m67.pdf}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $0.75~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in teal.
The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the orange dotted lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in teal, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more narrow and clustered than the \tgas\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}

\subsection{Distances from Asteroseismology}
\edit1{As a second comparison of these photometric parallax inferences, we compare to the inferred distances to the stars from asteroseismology.  Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars than the photometric parallax method. There are about $2200$ stars that overlap our data set here and the asteroseimic data from \citealt{huber17}. In \figurename~\ref{fig:seismic} we compare the estimates in the parallax space. We chose the parallax space, rather than the distance space, since the asteroseismic distance estimates are more precise, making it a less skewed inverted pdf. We compare the cumulative difference in the parallax estimates in units of mas; the same units as the precision of \tgas. The comparison shows that our photo parallax estimates are more similar to the asteroseismic estimates than the \tgas\ likelihoods. The left panel and the middle panel show the one to one comparison of asteroseismic parallax estimates to the \tgas\ likelihoods (in black) and the photo parallax posterior pdf (in blue) respectively for all $2200$ stars. The right panel shows the cumulative distribution function of the differences between the asteroseismic estimates and the \tgas\ likelihoods (in black) or the photoparallax estimates (in blue) for each star. The comparison to the photo parallax estimates, represented by the blue line, shows a distribution with smaller differences, and therefore represents estimates that are more similar to the asteroseismic estimates.}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{photoVSseismic.pdf}
\caption{\edit1{Proof of concept: parallax estimate comparisons. All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate (in black) or our photo parallax estimate (in blue). And the figure is meant to guide visual comparisons between \tgas\ and photo parallax, and therefore highlight differences in these estimates. The first panel shows the one to one comparison of the \tgas\ parallax estimates and the inverted distance estimates from \citealt{huber17} using asteroseismology, and the associated $1\sigma$ uncertainties. The orange line shows the one to one line. The second panel is similar, but instead of the \tgas\ parallax estimates, it visualizes our photo parallax estimates. This shows the shrinkage of the precision of the parallax estimates, and that the expectation values are more in agreement with the asteroseismic estimates. The third panel panel visualizes the difference between the asteroseismic estimate and a \gaia\ estimate, \tgas\ in grey and photo parallax in blue, as a cumulative distribution function of the difference in units of mas. The CDF of the photo parallax estimates is shifted to smaller differences, so the photoparallax estimates are therefore more similar to the asteroseismic estimates than \tgas\ likelihoods alone.}}

@@

\section{Discussion and Conclusions}

@@

\deleted{ Conclusion -- 
We forgo an explicit conclusion; we attempt to summarize the conclusions in the Abstract above.}
